---
title: "Preprocessing"
author: "Aishwarya"
date: "10/16/2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Loading all required packages

```{r tidying data}


library(tidyverse)
library(dplyr)
library(readr)
library(ggplot2)
library(lubridate)
library(tidyr)
#install.packages("ggmap")
library(ggmap)
library(caret)
library(modelr)

#install.packages('installr')
library(installr)

## install gganimate from GitHub using following code
# install.packages("devtools")
library(devtools)
#install_github("dgrtwo/gganimate")

## install ImageMagick package
#library(installr)
#install.ImageMagick()
#install.packages("gifski")

library(tibble)
library(lubridate)
library(maps)
library(ggthemes)
library(gganimate)
library(animation)

## FOR TEXT MINING , install wordcloud2 package from github
require(devtools)
# install_github("lchiffon/wordcloud2")

#install.packages("treemap")
#install.packages("leaflet")

library(treemap)
library(leaflet)


```
Using Google ggmap package for obtaining geospatial data of locations and storing it in 
 file "location_df"
use following commented code for getting location data for locations in 
employee reviews dataset from google API.

```{r location data, echo=FALSE}



#register_google('Your-google-API-Key')

#Address <- read_data %>% filter(is.na(location) == FALSE) %>% distinct(data, location)

#Address <- as.data.frame(Address)

#locations_df <-  mutate_geocode(Address, location)

#locations_df <- as_tibble(locations_df)

#write.csv(locations_df, "your_file_path_to_store")


```



importing and tidying dataset

```{r}

read_data <- read_csv("E:/classes resources/DS 5110/project/employee_reviews.csv" , 
                      guess_max = 1000 ,
                      na = c("none", "n/a") )
locations_df <- read.csv("E:/classes resources/DS 5110/project/locations_df.csv") %>% 
                as.tibble()


tidy_data <- read_data %>% full_join(locations_df) %>%
            select(-X) %>%
            separate(`job-title` , into=c("job-title","designation"), sep = "-") %>%
            mutate(dates = mdy(`dates`))





```



```{r}
# knowing number of employees in company 
tidy_data %>% 
  group_by(company) %>%
  count()


# current and former employees in each company
tidy_data %>% 
  group_by(company, `job-title`) %>%
  count()
```



```{r data summary for each observation}



data_intro <- tidy_data 


#Each company positive and negative overall ratings on map
data_intro %>% 
  filter(`company` != "NA",`overall-ratings` != "NA") %>%  
  mutate(`overall rating` = ifelse(`overall-ratings` > 3 , "Positive" , "Negative")) %>%
  ggplot() +
  borders("world", colour = "white", fill = "gray50")+ 
  geom_point(aes(x = lon, y = lat, colour =`overall rating`)) +
  scale_color_manual(values = c("red", "green")) + facet_wrap(~company) +theme_void()

########Interactive maps to visualise observations data.
library(leaflet)

###############Tree Map of jobtitles
library(treemap)
#install.packages("treemap")
title_count <- data_intro %>% group_by(designation) %>% summarize(count = n())%>% arrange(desc(count))

Top_title_count <- title_count[-1,] 

Top_20_title_count <- Top_title_count[1:20,]

treemap(Top_20_title_count,
        index = "designation",
        vSize = "count",
        palette = "RdBu",
        fontsize.title = 10 )


leaflet(Top_title_count) %>% addTiles() %>%
  addMarkers(lng = ~data_intro$lon, 
             lat = ~data_intro$lat,
             clusterOptions = markerClusterOptions(),
             
             popup = paste("<b>", "Company :", "</b>", data_intro$company, "<br>",
                           "<b>", "Employment :","</b>", data_intro$`job-title`, "<br>",
                           "<b>", "Designation :","</b>" ,data_intro$designation, "<br>",
                           "<b>", "Overall rating :","</b>" ,data_intro$`overall-ratings`))




```

Location vs overall rating

```{r location vs overall rating }

location_plot_data <- tidy_data


#Removing the rows which has 'none' as there rating and location
location_plot_data <- location_plot_data %>%
  filter(location != 'NA' )
  
#Top Locations to work for in Google
google <- location_plot_data %>%
  filter(company == "google") %>%
  group_by(location, lon, lat) %>%
  summarise(Mean = mean(`overall-ratings`), Count = n()) %>% 
  arrange((Count)) %>%
  filter(Count > 41) %>% #1% of total employees (4148)
  arrange(desc(Mean))
  
#Selecting top 5 locations for Google
googleTop5 <- google[1:5,]

#Bar plot for Google
ggplot(googleTop5, aes(x = reorder(location, Mean), y = Mean)) +
  geom_bar(stat = "identity", fill="green4") +
  labs(title="Top 5 Locations to work in Google" , x = "Top Locations" , y="Mean of Overall Rating") +
  coord_flip()

#World map plot for Google
ggplot(googleTop5, aes(x=lon, y=lat, color=location)) +
  borders("world", colour = "gray50", fill = "gray50") + 
  labs(title="Top 5 Locations to work in Google" , x = "Longitude" , y="Latitude") +
  geom_point()





```


AMAZON

```{r}
#Top Locations to work for in Amazon
amazon <- location_plot_data %>%
  filter(company == "amazon") %>%
  group_by(location, lon, lat) %>%
  summarise(Mean = mean(`overall-ratings`), Count = n()) %>% 
  arrange((Count)) %>%
  filter(Count > 140) %>% # approx less than 1% of total employees (16796)
  arrange(desc(Mean))

#Selecting top 5 locations for Amazon
amazonTop5 <- amazon[1:5,]

#Bar plot for Amazon
ggplot(amazonTop5, aes(x = reorder(location, Mean), y = Mean)) +
  geom_bar(stat = "identity", fill="gold") +
  labs(title="Top 5 Locations to work in Amazon" , x = "Top Locations" , y="Mean of Overall Rating") +
  coord_flip()

#World map plot for Amazon
ggplot(amazonTop5, aes(x=lon, y=lat, color=location)) +
  borders("world", colour = "gray50", fill = "gray50") + 
  labs(title="Top 5 Locations to work in Amazon" , x = "Longitude" , y="Latitude") +
  geom_point()


```


APPLE

```{r}
#Top Locations to work for in Apple
apple <- location_plot_data %>%
  filter(company == "apple") %>%
  group_by(location, lon, lat) %>%
  summarise(Mean = mean(`overall-ratings`), Count = n()) %>% 
  arrange((Count)) %>%
  filter(Count > 75) %>% #1% of total employees (7573)
  arrange(desc(Mean))

#Selecting top 5 locations for Apple
appleTop5 <- apple[1:5,]

#Bar plot for Apple
ggplot(appleTop5, aes(x = reorder(location, Mean), y = Mean)) +
  geom_bar(stat = "identity", fill="black") +
  labs(title="Top 5 Locations to work in Apple" , x = "Top Locations" , y="Mean of Overall Rating") +
  coord_flip()

#World map plot for Apple
ggplot(appleTop5, aes(x=lon, y=lat, color=location)) +
  borders("world", colour = "gray50", fill = "gray50") + 
  labs(title="Top 5 Locations to work in Apple" , x = "Longitude" , y="Latitude") +
  geom_point()


```


FACEBOOK

```{r}
#Top Locations to work for in Facebook
facebook <- location_plot_data %>%
  filter(company == "facebook") %>%
  group_by(location, lon, lat) %>%
  summarise(Mean = mean(`overall-ratings`), Count = n()) %>% 
  arrange((Count)) %>%
  filter(Count > 15) %>% #1% of total employees (1590)
  arrange(desc(Mean))

#Selecting top 5 locations for Facebook
facebookTop5 <- facebook[1:5,]

#Bar plot for Facebook
ggplot(facebookTop5, aes(x = reorder(location, Mean), y = Mean)) +
  geom_bar(stat = "identity", fill="blue") +
  labs(title="Top 5 Locations to work in Facebook" , x = "Top Locations" , y="Mean of Overall Rating") +
  coord_flip()

#World map plot for Facebook
ggplot(facebookTop5, aes(x=lon, y=lat, color=location)) +
  borders("world", colour = "gray50", fill = "gray50") + 
  labs(title="Top 5 Locations to work in Facebook" , x = "Longitude" , y="Latitude") +
  geom_point()




```

MICROSOFT

```{r}
#Top Locations to work for in Microsoft
microsoft <- location_plot_data %>%
  filter(company == "microsoft") %>%
  group_by(location, lon, lat) %>%
  summarise(Mean = mean(`overall-ratings`), Count = n()) %>% 
  arrange((Count)) %>%
  filter(Count > 110) %>% #1% of total employees (11735)
  arrange(desc(Mean))

#Selecting top 5 locations for Microsoft
microsoftTop5 <- microsoft[1:5,]

#Bar plot for Microsoft
ggplot(microsoftTop5, aes(x = reorder(location, Mean), y = Mean)) +
  geom_bar(stat = "identity", fill="orange3") +
  labs(title="Top 5 Locations to work in Microsoft" , x = "Top Locations" , y="Mean of Overall Rating") +
  coord_flip()

#World map plot for Microsoft
ggplot(microsoftTop5, aes(x=lon, y=lat, color=location)) +
  borders("world", colour = "gray50", fill = "gray50") + 
  labs(title="Top 5 Locations to work in Microsoft" , x = "Longitude" , y="Latitude") +
  geom_point()
```


NETFLIX

```{r}

#Top Locations to work for in Netflix
netflixTop5 <- location_plot_data %>%
  filter(company == "netflix") %>%
  group_by(location, lon, lat) %>%
  summarise(Mean = mean(`overall-ratings`), Count = n()) %>% 
  arrange((Count)) %>%
  filter(Count > 5) %>% #1% of total employees (562)
  arrange(desc(Mean))

#Selecting top 5 locations for Netflix
netflixTop5 <- netflixTop5[1:5,]

#Bar plot for Netflix
ggplot(netflixTop5, aes(x = reorder(location, Mean), y = Mean)) +
  geom_bar(stat = "identity", fill="red") +
  labs(title="Top 5 Locations to work in Netflix" , x = "Top Locations" , y="Mean of Overall Rating") +
  coord_flip()

#World map plot for Netflix
ggplot(netflixTop5, aes(x=lon, y=lat, color=location)) +
  borders("world", colour = "gray50", fill = "gray50") + 
  labs(title="Top 5 Locations to work in Netflix" , x = "Longitude" , y="Latitude") +
  geom_point()




```



YEAR VS OVERALL RATING


```{r}
year_read <- tidy_data

year_plot2 <- year_read %>% 
        mutate(dates = as.character(year_read$dates)) %>%
        mutate(dates = ifelse(is.na(dates), 0 ,dates)) %>%
        filter(dates != 0) %>%
        mutate(year = as.character(year(dates)))

# smooth curve to show yearly variations of overall ratings for each company
year_plot2 %>%
  filter(`company` != "NA") %>%
  ggplot(aes(x= as.numeric(`year`),y = `overall-ratings`)) +
  geom_smooth(aes( colour=`company`),se=FALSE) +
  facet_wrap( ~ `company`)+
  labs(title= "Yearly average ratings for all companies" , x = "year" , y="overall rating") +
scale_x_continuous(breaks = c(2008, 2009, 2010, 2011, 2012, 2013, 2014, 2015, 2016, 2017, 2018))+
  theme(axis.text.x = element_text(angle=45 ,hjust =1))
```





```{r}
year_plot1 <- year_read %>% 
        mutate(dates = as.character(year_read$dates)) %>%
        mutate(dates = ifelse(is.na(dates), 0 ,dates)) %>%
        filter(dates != 0) %>%
        mutate(year = as.character(year(dates)))

# yearly statistics of overall ratings for each company
year_plot1 %>%
  filter(`company` != "NA") %>%
  filter(`year` %in% c("2010":"2018")) %>%
  ggplot() +
  geom_boxplot(aes(x=`year`, y =`overall-ratings`, fill = `company`), position = "dodge2")+
  facet_wrap( ~ `company`) +
  theme(axis.text.x = element_text(angle=45 ,hjust =1)) +
  labs(title= "statistics of ratings yearly for all companies" , x = "year" , y="overall rating")


# plotted through manually calculating the means for each year
year_plot2 %>%
  filter(`company` != "NA") %>%
  group_by(`company`,`year`) %>%
  summarize(mean_overall = mean(`overall-ratings`)) %>%
  ggplot(aes(x= as.numeric(`year`),y = `mean_overall`)) +
  geom_point(aes(colour=`company`)) +
  geom_line(aes(colour=`company`)) +
  scale_x_continuous(breaks = c(2008, 2009, 2010, 2011, 2012, 2013, 2014, 2015, 2016, 2017, 2018))+
  facet_wrap( ~ `company`)+
  labs(title= "Yearly average ratings for all companies" , x = "year" , y="mean overall rating") +
  theme(axis.text.x = element_text(angle=45 ,hjust =1))
```


Animated plots for yearly plots 

```{r}
year_animated <- year_plot2 %>%
  filter(`company` != "NA") %>%
  group_by(`company`,`year`,`lon`,`lat`) %>%
  summarize(mean_overall = mean(`overall-ratings`))
year_animated

year_animated_amazon <- year_animated %>%
                        filter(company == "amazon")

year_animated_apple <- year_animated %>%
                        filter(company == "apple")

year_animated_facebook <- year_animated %>%
                        filter(company == "facebook")

year_animated_google <- year_animated %>%
                        filter(company == "google")
year_animated_microsoft <- year_animated %>%
                        filter(company == "microsoft")
year_animated_netflix <- year_animated %>%
                        filter(company == "netflix")


world <- ggplot() +
  borders("world", colour = "black", fill = "gray80") +
  borders("state") +
  theme_map()
```

yearly animated plot for Amazon

```{r}
year_map2_amazon <- world +
  geom_point(year_animated_amazon, 
             mapping=aes(x = lon, y = lat, colour = cut_width(`mean_overall`, 1, boundary = 0)),
             size= 5 , alpha = 0.5) +
  coord_cartesian(xlim = c(-60,-140) , y=c(25,50)) +
  transition_manual(`year`) +
  labs(title = "Amazon , year : {current_frame}") +
  scale_colour_manual(values = c("blue","orange","red","green") , name = "ratings") 


year_map2_amazon <- animate(year_map2_amazon, fps = 20, duration = 10)

```

yearly animated plot for Apple 
```{r}
year_map2_apple  <- world +
  geom_point(year_animated_apple, 
             mapping=aes(x = lon, y = lat, colour = cut_width(`mean_overall`, 1, boundary = 0)), 
             size = 5 ,alpha = 0.5) +
  coord_cartesian(xlim = c(-60,-140) , y=c(25,50)) +
  transition_manual(`year`) +
  labs(title = "Apple, year : {current_frame}") +
  scale_colour_manual(values = c("blue","orange","red","green"), name = "ratings") 

year_map2_apple <- animate(year_map2_apple, fps = 20, duration = 10)
```

yearly animated plot for Facebook 

```{r}
year_map2_facebook <- world +
  geom_point(year_animated_facebook,
             mapping=aes(x = lon, y = lat, colour = cut_width(`mean_overall`, 1, boundary = 0)), 
             size=5, alpha = 0.5) +
  coord_cartesian(xlim = c(-60,-140) , y=c(25,50)) +
  transition_manual(`year`) +
  labs(title = "Facebook ,year : {current_frame}") +
  scale_colour_manual(values = c("blue","orange","red","green"), name = "ratings") 
  
animate(year_map2_facebook, fps = 20, duration = 10)
```

yearly animated plot for google 

```{r}
year_map2_google <- world +
  geom_point(year_animated_google,
             mapping=aes(x = lon, y = lat, colour = cut_width(`mean_overall`, 1, boundary = 0)),
             size= 5 ,alpha = 0.5) +
  coord_cartesian(xlim = c(-60,-140) , y=c(25,50)) +
  transition_manual(`year`) +
  labs(title = "Google, year : {current_frame}") +
  scale_colour_manual(values = c("blue","orange","red","green"), name = "ratings") 

year_map2_google <- animate(year_map2_google, fps = 20, duration = 10)
```

 yearly animated plot for Microsoft
```{r}
year_map2_microsoft <- world +
  geom_point(year_animated_microsoft,
             mapping=aes(x = lon, y = lat, colour = cut_width(`mean_overall`, 1, boundary = 0)), 
             size= 5, alpha = 0.5) +
  coord_cartesian(xlim = c(-60,-140) , y=c(25,50)) +
  transition_manual(`year`) +
  labs(title = "Microsoft, year : {current_frame}") +
  scale_colour_manual(values = c("blue","orange","red","green"), name = "ratings") 

year_map2_microsoft <- animate(year_map2_microsoft, fps = 20, duration = 10)
```

yearly animated plot for Netlfix
```{r}

year_map2_netflix <- world +
  geom_point(year_animated_netflix,
             mapping=aes(x = lon, y = lat, colour = cut_width(`mean_overall`, 1, boundary = 0)),
             size = 5, alpha = 0.5) +
  coord_cartesian(xlim = c(-60,-140) , y=c(25,50)) +
  transition_manual(`year`) +
  labs(title = "Netflix, year : {current_frame}") +
  scale_colour_manual(values = c("blue","orange","red","green"), name = "ratings") 
year_map2_netflix <- animate(year_map2_netflix, fps = 20, duration = 10)
anim_save("netflix1.gif", year_map2_netflix ) 

year_map2_world <- world +
  geom_point(year_animated,
             mapping=aes(x = lon, y = lat, size = `mean_overall`,
                                        colour= company),  alpha = 0.5) +
  transition_manual(`year`) +
  labs(title = "year : {current_frame}") +
  scale_size_continuous(range = c(0,5), breaks = c(1,2,3,4,5)) +
  scale_colour_manual(values = c("deepskyblue","coral","darkorange1","red","blue","green"), name = "ratings") 
year_map2_world <- animate(year_map2_world, fps = 20, duration = 10)


```

overall ratings vs other aspects


```{r}
general_data <- tidy_data

 general_data <- general_data %>% 
   filter(`overall-ratings` !="none" & `culture-values-stars`!="none" & `work-balance-stars`!="none" &`carrer-opportunities-stars` !="none" &`comp-benefit-stars`!="none" & `senior-mangemnet-stars`!="none")
 
companynames <- general_data %>% 
  group_by(company) %>% 
  summarise(observations = n())

company_overallrating <-general_data %>%
  group_by(`overall-ratings`,company) %>% 
  summarise(rating = n())
```

overall rating proportions each company
```{r}

full_join(companynames,company_overallrating) %>% 
  mutate(fine=rating/observations) %>% 
  ggplot()+
  geom_bar(mapping = aes(x=company,y=fine,fill=as.factor( `overall-ratings`)),
           stat = "identity",position = "dodge")+
  labs(title="Proportion of overall rating vs Company", x="Company",
       y="Proportion of overall rating",fill="overall-ratings")


company_workbalance <- general_data  %>% 
  group_by(`work-balance-stars`,company) %>%
  summarise(total=n())
full_join(company_workbalance,companynames) %>%
  mutate(workbalance= total/observations ) %>%
  ggplot()+geom_bar(mapping = aes(x=company,y=workbalance,fill=as.factor(`work-balance-stars`)),
                    stat = "identity",position = "dodge")+
  labs(title="Proportion of Work-Balance-Stars vs Company", x="Company",
       y="Proportion of Work-Balance-Stars",fill="work-balance-stars")


general_data[["dates"]] <- mdy(general_data[["dates"]])
general_data <- mutate(general_data,year=year(dates))
general_data <- mutate(general_data,pos= str_sub(general_data[["job-title"]],1,16))
general_data <- mutate(general_data,pos=str_trim(general_data[["pos"]]))

company_position <-general_data %>% 
  group_by(pos,company) %>% 
  summarise(position=n())

 full_join(company_position,companynames) %>%
   mutate(probability=position/observations) %>% 
   ggplot()+geom_bar(mapping = aes(x=company,y=probability,fill=pos),
                     stat = "identity",position = "dodge")+
   labs(title="Employee Status", x="Company", y="Proportion")


positions <- general_data %>% 
  group_by(company,pos) %>% 
  summarise(totalpos=n())

overall_pos <- general_data %>% 
  group_by(company,pos,`overall-ratings`) %>%
  summarise(eachpos=n())

full_join(positions,overall_pos) %>% mutate(prop=eachpos/totalpos) %>% 
  ggplot()+
  geom_bar(mapping = aes(x=pos,y=prop,fill=as.factor(`overall-ratings`),
                         position = "dodge"),stat = "identity",position = "dodge")+
  labs(title="Proportion of overall rating vs Employee Status", 
       x="Employee Status", y="Proportion of overall rating",fill="overall-ratings")+
  facet_wrap(~company)+coord_flip()



average<- general_data %>% 
  gather(`overall-ratings`,`work-balance-stars`,`culture-values-stars`,
         `carrer-opportunities-stars`,`comp-benefit-stars`,`senior-mangemnet-stars`,
         key="fields",value = "ratings") %>% group_by(company,fields,ratings) %>% 
  summarise(totals=n())%>% mutate(prod=as.numeric(ratings)*totals) %>% 
  group_by(company,fields) %>% summarise(totalsum=sum(prod))



full_join(average,companynames) %>% group_by(company,fields) %>%
  mutate(totalratings=totalsum/observations) %>% 
  ggplot()+geom_bar(mapping = aes(x=company,y=totalratings,fill=company,position = "dodge"),
                    stat = "identity") +facet_wrap(~fields)+
  coord_flip()+
  labs(title="Company vs Average of all six-ratings", y="Average of all six-ratings",
       x="company",fill="company") + 
  scale_fill_manual(values = c("gold","black","blue","green4","orange3","red"))


average_pos<- general_data %>%
  gather(`overall-ratings`,`work-balance-stars`,`culture-values-stars`,
         `carrer-opportunities-stars`,`comp-benefit-stars`,`senior-mangemnet-stars`,
         key="fields",value = "ratings") %>% group_by(company,fields,ratings,pos) %>%
  summarise(posavg=n())%>% mutate(product=as.numeric(ratings)*posavg) %>%
  group_by(company,fields,pos) %>% summarise(numerator=sum(product))



full_join(companynames,average_pos)  %>% 
  mutate(totalratings=numerator/observations) %>% 
  ggplot()+geom_bar(mapping = aes(x=company,y=totalratings,fill=pos),
                    stat = "identity",position = "dodge") +facet_wrap(~fields)+
            coord_flip()+
  labs(title="Company vs Proportion of Average ratings contributed by Current asnd Former Employees",
       x="Company", y="Average of all six-ratings",fill="position")





```

